---
title: "Watershed Regression - Part 2: Censored MCMC Generalized Linear Mixed Models"
author: "Christian Nilsen"
output:
  html_document:
    toc: yes
    highlight: zenburn
    df_print: paged
  html_notebook:
    toc: yes
    highlight: zenburn
    df_print: paged
    fig_caption: yes
---

# License

> Copyright (c) 2020 Geosyntec Consultants, Inc. 
[Mozilla Public License Version 2.0](https://choosealicense.com/licenses/mpl-2.0/)


This software is provided "as is", without warranty of any kind, express or implied, including but not limited to the warranties of merchantability, fitness for a particular purpose and noninfringement. In no event shall the authors or copyright holders be liable for any claim, damages or other liability, whether in an action of contract, tort or otherwise, arising from, out of or in connection with the software or the use or other dealings in the software.  




# Description 

This notebook is the second step in the watershed regression task. It uses the relationships from Part 1 (linear models) to run generalized linear mixed models with censored dependent variables. This approach uses  Markov chain Monte Carlo techniques and relies heavily on the ```MCMCglmm``` package (Hadfield 2009).  


Here are the steps in the code below:
   
1. Code set up  
2. Get and prepare monitoring data  
3. Merge spatial data with monitoring data  
4. Remove mulitcolinear predictors   
5. Peform model selection   
6. Select the best model (these results are then used in the Bayesian model in step 2)  
  


# Setup 
clear workspace and load libraries. (code not shown in html file ) 

```{r knitr_init, cache=FALSE, include=FALSE}
rm(list = ls())
library(knitr)
knitr::include_graphics
library(rmdformats)
library(tibble)
library(tidyverse)
library(ggplot2)
library(caret)
library(psych)
library(DataExplorer)
library(dplyr)
library(car)
library(lme4)
library(nlme)
library(hrbrthemes)
library(sjPlot)
#library(ggplot2); 
library(MCMCglmm)
library(plotMCMC)
library(modelsummary)
#library(lmec)
#library(Metrics)
#library(ggstatsplot)
#install.packages('modelsummary')
#library(modelsummary)
#library('emmeans')
#library('tidybayes')
theme_set(theme_ipsum_rc())

#detach("package:datasets", unload=TRUE)
## Global options
options(max.print="75")
opts_chunk$set(cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)

set.seed(50)
```
```{r}

#library("renv")
library(caret)
library(readr)
library(dplyr)
 
```


```{r}
setwd("~/repos/stormwaterheatmap/R-scripts/WatershedRegression")



```



# Get and prepare monitoring data

The stormwater outfall data is available from the Department of Ecology at https://data.wa.gov/Natural-Resources-Environment/Municipal-Stormwater-Permit-Outfall-Data/d958-q2ci.

A .csv file is saved in ```WatershedRegression/data/S8_data.csv``` 

```{r echo=TRUE}
all.S8.data <- read_csv("data/S8_data.csv", 
    col_types = cols(field_collection_end_date = col_date(format = "%m/%d/%Y"), 
        field_collection_start_date = col_date(format = "%m/%d/%Y")) )

#filter out rejected data
all.S8.data <- (filter(all.S8.data,!result_data_qualifier %in% 'REJ'))

#filter out replicates 
all.S8.data <- (filter(all.S8.data,!sample_replicate_flag %in% 'Y'))

#change nondetect warnings to detects
warnings <- all.S8.data$nondetect_flag == "WARNING"
all.S8.data$nondetect_flag[warnings] <- FALSE 

#Change NA to detect
all.S8.data$nondetect_flag[is.na(all.S8.data$nondetect_flag)] <- FALSE

#Change season to factor 
all.S8.data$season <- as.factor(all.S8.data$season)


```

The chunk below makes a list of parameters we are interested in. 

```{r Select Parameters: }
#Select Parameters
params <- c('Zinc - Water - Total',
 'Copper - Water - Total',
 'Nitrite-Nitrate - Water - Dissolved',
 'Lead - Water - Total',
 'Total Phosphorus - Water - Total',
 'Total Suspended Solids - Water - Total',
 'Total Phthalate - Water - Total',
'Total PAH - Water - Total',
#'Chrysene - Water - Total',
'CPAH - Water - Total',
'HPAH - Water - Total' 
#'Total Kjeldahl Nitrogen - Water - Total',
#'Total PCB - Water - Total'
)

#save a list of all the parameters in case we want to use mor. 
params.all <- data.frame(unique(all.S8.data$parameter))
s8data <- all.S8.data 
#
```

Clean up extracted data and rename columns: 

```{r }
s8data <- all.S8.data %>% 

  dplyr::select(
    study_name,
    location_id,parameter,
    type,
    season,
    new_result_value,
    nondetect_flag,
    study_id,
    access_id,
    field_collection_end_date,
    field_collection_start_date,
    type)


#rename some columns
colnames(s8data)[colnames(s8data) == "location_id"] <- "Location"
colnames(s8data)[colnames(s8data) == "new_result_value"] <-
  "concentration"
s8data$nondetect_flag <- as.logical(s8data$nondetect_flag)
s8data$concentration <- as.numeric(s8data$concentration)

```

## Check for outliers

Set up a jitter plot of all the data to look for outliers: 
  
```{r fig.height=8}

scatter_cocs <- function(df.coc,title) {
 p <- ggplot(df.coc, aes(1, concentration)) + geom_jitter() + labs(
  title = title,
  subtitle = "Data collected 2009-2013",
  caption =
    " Data source: Ecology, 2015",
  x = "Observations"
  #y = y.title
)  
p + facet_wrap( ~ parameter, ncol=3, scales = 'free') 
}

scatter_cocs(s8data[which(s8data$parameter %in% params),],'All Observations')
```
   
outliers are apprent for TSS, TP, and no2-no3. Remove these. 
```{r remove outliers, fig.height=8}
#remove and replot 

outlierParams <- c("Total Suspended Solids - Water - Total", "Total Phosphorus - Water - Total", "Nitrite-Nitrate - Water - Dissolved")

#This removes the highest values 
outlierVals <-
  top_n(group_by(s8data[which(s8data$parameter %in% outlierParams), ], parameter), 1, concentration)$concentration

s8data <- s8data %>%
  group_by(parameter) %>%
  slice(which(!(
    parameter %in% outlierParams & concentration %in% outlierVals
  )))

scatter_cocs(s8data[which(s8data$parameter %in% params),],'All Observations - Outliers Removed')

```
   
Looks better. Move on to the next Chunk. 

# Get Spatial data and merge  

```{r}

#Spatial predcitors have been extracted and saved as a csv file. 

spatial_data <- read_csv("C:/Users/cnilsen/Google Drive/repos/spatialPredictors.csv")
#RES and COM are compositional data. Change to a ratio
spatial_data$LU_ratio = spatial_data$COM/spatial_data$RES 
spatial_data <- dplyr::select(spatial_data, -c(RES,COM,.geo))
#merge spatial predictors with monitoring data 
s8data.wPredictors <<- merge(s8data, spatial_data)%>% 
  dplyr::select(-c(depSplusN))

# spatial_data<-read_csv("data/spatialPredictors42.csv", col_types = cols(X1 = col_skip()))
# 
# #merge spatial predictors with monitoring data 
# s8data.wPredictors <- merge(s8data, spatial_data)

```


# Perform MCMC modeling 
## Functions 
Some helper functions to help 

##### Function to add a survival object to the S8 dataframe 
```{r}
add_surv <- function(df) {
  df$cenMin <- ifelse(df$nondetect_flag,-Inf, (df$concentration))
  df$cenMax  <- (df$concentration)
  df$cenMin_log <- ifelse(df$nondetect_flag,-Inf, log(df$concentration))
  df$cenMax_log  <- log(df$concentration)
  
  return(df)
}
```
##### Function to return a chart of predictions from the model 
```{r}
scatter_predict <- function(model_df,predictions) {
# model_to_predict <- CuModel 
# coc = params[2]
# df <- (subset(s8data.wPredictors, parameter == coc)) %>%
#   add_surv()
# predictions <- predict(model_to_predict, newdata=df, 
#          type="response", interval="none", level=0.95, it=NULL, 
#          posterior="all", verbose=FALSE, approx="numerical")
# 
 obs <- log(model_df$concentration)
 
 ggstatsplot::ggscatterstats(
  data =tibble(p = predictions, obs = obs,L = model_df$Location),
  x = p,
  y = obs,
  type = "bf",
  point.width.jitter = 0.02,
  #point.height.jitter = 0.1,
  marginal = FALSE,
  xlab = "Predicted log(µg/L)",
  ylab = "Observed  log(µg/L)",
  title = coc,
  results.subtitle = FALSE,
  subtitle = "Predictions vs. Observations",
  smooth.line.args = list(size = 1, color = "blue"),
  messages = FALSE
)
}

```

##### add tidy and glance functions to the mcmcglmm objects since they don't play nicely with tidyverse objects. 
```{r}
# add custom functions to extract estimates (tidy) and goodness-of-fit (glance) information
tidy.MCMCglmm <- function(object, ...) {
    s <- summary(object, ...)
    ret <- tibble::tibble(term = row.names(s$solutions),
                          estimate = s$solutions[, 1],
                          conf.low = s$solutions[, 2],
                          conf.high = s$solutions[, 3])
    ret
}
glance.MCMCglmm <- function(object, ...) {
    ret <- tibble::tibble(dic = object$DIC,
                          n = nrow(object$X))
    ret
}

# estimate a simple model
#model <- MCMCglmm(PO ~ 1 + plate, random = ~ FSfamily, data = PlodiaPO, verbose=FALSE, pr=TRUE)
```

## mcmc_calc function  

This is the main funciton to run the mcmc model. It does the following:   
1. subsets a dataframe to include only the parameter we want to predict  
2. adds a survival object to handle censored data  
3. sets up a simiple prior structure  
4. performs mcmc modeling on either the log-transformed or non-log transformed responses.   
5. returns the results   

```{r}
mcmc_calc <- function(coc.local, fixed_list, lhs) {
  
  df <- (subset(s8data.wPredictors, parameter == coc.local))

   data <-
  df %>%
  add_surv()
#make the prior_structures 
prior.1<-list(R=list(V=1, fix=1), G=list(G1=list(V=1, nu=0.002)))
prior.2<-list(R=list(V=2, fix=1), G=list(G1=list(V=1, nu=0)))

  
  if (lhs == 'log') {
    mcmc_formula <-
      as.formula(paste(
        "cbind(cenMin_log, cenMax_log) ~ ",
        paste0(fixed_list , collapse = "+")
      ))
  }
  else {
    mcmc_formula <-
      as.formula(paste(
        "cbind(cenMin, cenMax) ~ ",
        paste0(fixed_list , collapse = "+")
      ))
  }
  
  mcmc_results <-
    MCMCglmm(
      mcmc_formula,
      random = ~ Location,
      data =  data,
      family = "cengaussian", 
      verbose = FALSE, prior = prior.1, singular.ok = TRUE,
      nitt = 60000, thin = 13, burnin = 10000
    )
    return((mcmc_results))
}
```



##### function that returns bayesian plots
```{r}
# Do some predictive checks 
library(bayesplot)
color_scheme_set("blue")
bayes_plots <- function(fit,coc,df) {
#fit <- TSSModel

#coc = 'Total Suspended Solids - Water - Total'

#df <- (subset(s8data.wPredictors, parameter == coc)) %>%
 # add_surv()
yrep_c <- predict(fit, newdata=df, 
         type="response", interval="confidence", level=0.9, it=NULL, 
         posterior="all", verbose=FALSE, approx="numerical")
yrep_p <- predict(fit, newdata=df, 
         type="response", interval="prediction", level=0.9, it=NULL, 
         posterior="all", verbose=FALSE, approx="numerical")

#show uncertainty intervals under esimated posterior density curves 
plot.1 <- mcmc_areas(fit$Sol,prob = 0.80, pprob_outer = 0.95,point_est="mean")+ggplot2::labs(title = coc, subtitle   = "Posterior distributions with medians and 80% intervals")

#generate scatter plot of predictions 

colnames(yrep_p) <-  c("fit.p", "lwr.p", "upr.p")
scatterdata <- cbind(df, yrep_c, yrep_p)

#generate scatter plot of predictions 
plot.2 <- ggplot(scatterdata) + 
  geom_ribbon(aes(ymin = lwr.p, ymax = upr.p, x = fit),fill="grey", alpha = 0.5) + 
  geom_ribbon(aes(ymin = lwr, ymax = upr, x = fit), fill = "grey", alpha = 0.8) + 
  geom_line(aes(x=fit,y=fit),color="blue",linetype=5)+
  geom_point(aes(x = fit, y = log(concentration)), alpha = 0.5)+
  ggplot2::labs(x="yrep",y="fit",title = coc, subtitle   = "Scatter plot of observed data vs simulated",caption="dark shade: confidence intervals \n light shade: prediction intervals")


#simulate with 100 draws  
ysim <- (simulate(fit,nsim = 100))


#overlay of predictions 
plot.3<- ppc_dens_overlay(log(df$concentration),t(ysim))+ggplot2::labs(x="log concentration, μg/L ",title = coc, subtitle   = "Observed (y) vs. simulated draws (yrep)")


return(list(plot.1,plot.2,plot.3))
}
```

```{r}
#function for other diagnostic plots 
diagnostic_plots <- function(chains,coc) {
  plotTrace(chains,axes=TRUE,same.limits=TRUE)
  plotDens(chains,main=paste('Posterior Distributions \n',coc),probs=c(0.050,0.950),same.limits=FALSE)}
```


# Run model 

For each chunk below, we run the model and output diagnostic and prediction plots. 

## Zinc   
```{r fig.height=7}
# 
coc <- params[1]
ZnModel <- mcmc_calc(coc,c('impervious'),'log')
mod <- ZnModel
summary(mod)
df <- (subset(s8data.wPredictors, parameter == coc)) %>%
  add_surv()
bayes_plots(mod,coc,df)


```

## Copper    

```{r fig.height=7}
coc = 'Copper - Water - Total'
CuModel <- mcmc_calc(coc,c(
'rev_logTraffic','impervious'),'log')
mod <- CuModel
summary(mod$Sol)
summary(mod)
df <- (subset(s8data.wPredictors, parameter == coc)) %>%
  add_surv()
bayes_plots(mod,coc,df)

```



## Nitrite-Nitrate

```{r}
coc = 'Nitrite-Nitrate - Water - Dissolved'
NN_model <- mcmc_calc(coc,c(
'LU_ratio'),'log')
mod <- NN_model
summary(mod$Sol)
summary(mod)
df <- (subset(s8data.wPredictors, parameter == coc)) %>%
  add_surv()
bayes_plots(mod,coc,df)

```

## Lead   

```{r fig.height=7}
coc <- params[4]
PbModel <- mcmc_calc(params[4],c('impervious'),'log')

mod <- PbModel
summary(mod$Sol)
summary(mod)

df <- (subset(s8data.wPredictors, parameter == coc)) %>%
  add_surv()
bayes_plots(mod,coc,df)


```
## Cadium   

```{r fig.height=7}
coc <- 'Cadmium - Water - Total'
CdModel <- mcmc_calc('Cadmium - Water - Total',c('impervious','logPopulation'),'nonlog')
mod <- CdModel
summary(mod$Sol)
summary(mod)

df <- (subset(s8data.wPredictors, parameter == coc)) %>%
  add_surv()
bayes_plots(mod,coc,df)

```

## Total Phosphorus  

```{r fig.height=7}
coc = 'Total Phosphorus - Water - Total'
TPModel <- mcmc_calc('Total Phosphorus - Water - Total',c('rev_logTraffic'),'log')
mod <- TPModel
summary(mod$Sol)
summary(mod)

df <- (subset(s8data.wPredictors, parameter == coc)) %>%
  add_surv()
bayes_plots(mod,coc,df)
```

## Total Kjeldahl Nitrogen  

```{r fig.height=7}
coc = 'Total Kjeldahl Nitrogen - Water - Total'
TKNModel<- mcmc_calc('Total Kjeldahl Nitrogen - Water - Total','rev_logTraffic','log') 
mod <- TKNModel
df <- (subset(s8data.wPredictors, parameter == coc)) %>%
  add_surv()
bayes_plots(mod,coc,df)
```
## TSS  

```{r}
coc = 'Total Suspended Solids - Water - Total'
TSSModel <- mcmc_calc('Total Suspended Solids - Water - Total','rev_logTraffic','log')
df <- (subset(s8data.wPredictors, parameter == coc)) %>%
  add_surv()
bayes_plots(TSSModel,coc,df)


```
'Total PAH - Water - Total',
```{r}
coc = 'Total PAH - Water - Total'
PAHModel<- mcmc_calc('Total PAH - Water - Total','LU_ratio','log') 
mod <- PAHModel
df <- (subset(s8data.wPredictors, parameter == coc)) %>%
  add_surv()
bayes_plots(mod,coc,df)
```

'CPAH - Water - Total'
```{r}
coc = 'CPAH - Water - Total'
CPAHModel<- mcmc_calc('CPAH - Water - Total',c('logPopulation','rev_logTraffic'),'log') 
mod <- CPAHModel
df <- (subset(s8data.wPredictors, parameter == coc)) %>%
  add_surv()
bayes_plots(mod,coc,df)
```

'HPAH - Water - Total' 
```{r}
coc = 'HPAH - Water - Total'
HPAHModel<- mcmc_calc('HPAH - Water - Total',c('LU_ratio'),'log') 
mod <- HPAHModel
df <- (subset(s8data.wPredictors, parameter == coc)) %>%
  add_surv()
bayes_plots(mod,coc,df)
```


# Summary  

Summarize posterior results for use in heatmap. 

```{r}
#summarize 
metals <- list()
metals[['Total Copper']] <- (CuModel)
metals[['Total Zinc']] <- (ZnModel)
metals[['Total Cadmium']]<- (CdModel)
metals[['Total Lead']]<- (PbModel)

others <- list() 
others[['Total Phosphorus']] <- TPModel
others[['Total Kjeldahl Nitrogen']] <- TKNModel
others[['Total Suspended Sediment']] <-  TSSModel
#others['FC'] <- FCModel 





msummary(metals,title='Total Metals',statistic_vertical = TRUE,statistic = 'conf.int', conf_level = 0.95)
msummary(others,title='Nutrients and TSS',statistic_vertical = TRUE,statistic = 'conf.int', conf_level = 0.95)
```


Print out a messy list of results for the heatmap 
```{r}
sols <- list()
sols[['Copper']] <-summary(CuModel)$solutions  
sols[['Zinc']] <-summary(ZnModel)$solutions  
sols[['Cadmium']] <-summary(CdModel)$solutions  
sols[['Lead']] <-summary(PbModel)$solutions  
sols[['TP']] <-summary(TPModel)$solutions  
sols[['TKN']] <-summary(TKNModel)$solutions  
sols[['TSS']] <-summary(TSSModel)$solutions  

for (n in 1:length(sols)){
  print(kable(sols[[n]],row.names = TRUE, caption = paste(names(sols)[n]," - Summary of MCMC model")))
}
```




